Modeling tabular icebergs coupled to an ocean model

### 2017-03-31 JAMES_iceberg_only.tex
Another update

\authors{A. A. Stern\affil{1}, A. Adcroft\affil{1}, O. Sergienko\affil{1}, G. Marques\affil{1}, R. Hallberg\affil{1}}

@@

\begin{abstract}
Large tabular icebergs calved from Antarctic ice shelves have long lifetimes (due to their large size), during which they drift across large distances, altering ambient ocean circulation, bottom-water formation, sea-ice formation and biological primary productivity in the icebergs' vicinity.
However, despite their importance, the current generation of ocean circulation models do not represent large tabular icebergs.
In this study we develop a novel framework to model large tabular icebergs submerged in the ocean. In this framework, tabular icebergs are represented by Lagrangian elements that drift in the ocean, and are held together and interact with each other via bonds.
%We describe the Lagrangian bonded iceberg model and discuss a number of the technical elements involved in its development. 
%We demonstrate the capabilities of the model by modeling a tabular iceberg drifting away an idealized ice shelf.} 
A break of these bonds allows the model to emulate calving events (i.e. detachment of a tabular iceberg from an ice shelf) and tabular icebergs breaking up into smaller pieces. Idealized simulations of a calving tabular iceberg, its drift, and breaking into two pieces, demonstrate capabilities of the developed framework.
\end{abstract}

@@

In order to allow icebergs to travel large distances, we model the icebergs in a Lagrangian framework (as in the point particle iceberg drift models described above). However in our model icebergs are no longer treated as point particles that interact with the ocean at a single location. Instead icebergs are given physical structure, so that they interact with the ocean across multiple ocean grid cells, depress the ocean surface over a wide area, and can interact with other icebergs (Figure \ref{fig:Hex_schematic}). This is done by assigning a finite surface area and shape to the Lagrangian elements, which allows the elements to behave as if they have a finite extent. 
%Element interact with the ocean by distributing their weight, surface area and melt fluxes over multiple ocean grid cells in a way which is consistent with the shape of the ice element.
The finite extent of an element is transmitted by the ocean by distributing the element's weight, surface area and melt fluxes over multiple ocean grid cells in a way which is consistent with the shape of the ice element. 
Finite-extent elements interact with each other via repulsive forces which are applied when the boundaries of the elements overlap.
%Elements interact with other elements when the boundaries of the elements overlap (see schematic in Figure \ref{fig:Hex_schematic}).

Large tabular icebergs can then be represented by `bonding' together multiple ice elements into larger structures using numerical bonds (Figure \ref{fig:Hex_schematic}).
The numerical bonds hold the ice elements together and allow a collection of elements to move as a unit. 
This allows tabular icebergs to drift in the ocean when forced by ocean currents and wind.
%Figure \ref{fig:Hex_schematic}) shows a schematic of the difference between point-particle icebergs, finite extent icebergs and tabular icebergs constructed out of multiple elements. 
%Figure \ref{fig:Hex_schematic} shows a schematic of point-particle elements, elements with finite extent and bonded elements, and shows schematically how tabular icebergs can be constructed out of multiple elements.
An advantage of representing tabular icebergs using numerical bonds is that by breaking the bonds, we can simulate iceberg calving (e.g.: Figure \ref{fig:SST_Collapse}), or the response to an iceberg fracturing into multiple smaller pieces.

@@

%As mentioned above, we simplify the LBIM model by choosing to not explicitly calculate the angular momentum of the ice elements, which is computationally expensive. Instead, we allow rotational motion of larger structures emerge as a consequence of bond orientation and collective motion \citep{Jakobsen2001}.
%

@@

The Lagrangian Bonded Iceberg Model (LBIM) is a Lagrangian particle-based model in that the objects of the model are Lagrangian elements. Each element represents a column of ice that is floating in the ocean, and has a position, velocity, mass, and a set of dimensions, which can evolve in time. The motion of each element is determined by a momentum equation which is solved in the (Lagrangian) reference frame of the element. The elements experience oceanic and atmospheric forces, which are either prescribed, or computed by coupling the LBIM to an ocean/atmosphere model. The LBIM elements also interact with one another via attractive and repulsive interactive forces, and can be bonded together to form larger structures. The angular momentum of the elements is not modeled explicitly; instead rotational motion of larger structures emerge as a consequence of bond orientation and collective motion.

In different contexts, the LBIM elements can be thought to represent individual icebergs, sea ice flows, or, when the elements are bonded together, they can represent larger structures such as tabular icebergs or ice shelves.

%In this latter case, the LBIM can be categorized as a discrete element model (DEM). 
% in the sense that large structures are constructed from smaller independent Lagrangian elements.

The LBIM is developed on the code base of an existing iceberg drift model \citep{Martin2010, Stern2016}. When run with the correct set of runtime flags, the model runs as a traditional point-particle iceberg drift model.
%
%In this section, we discuss the equations of motion governing the element movement and the thermodynamic equations governing the decay of the elements. The also discuss the interactive forces

@@

% 
When ice elements move alone (without interactions with other elements), they can be thought of as representing individual (or clusters of) small icebergs, and follow the same equations described in the iceberg drift model of \citet{Martin2010} (based on the equations outlined in \citet{Bigg1997} and \citet{Gladstone2001}).
These environmental forces are provided for completeness in Appendix A.

In addition to the external forces, the elements in the LBIM experience interactive forces due to the presence of other elements. Two types of interactive forces are included between elements. The first force is a repulsive force which is applied to elements to prevent them from overlapping the boundaries of the neighboring elements. The second interactive force is a force due to numerical `bonds', and is only applied if two elements are labelled as `bonded'. When two elements are bonded, each element feels an attractive force that prevents the elements from moving too far apart from one another.  The details of the interactive forces are provided in below. 
%The interactive forces between two bonded elements are defined such that in the absence of other forces the elements come to rest adjacent to one another, with no overlap of the iceberg areas. \textcolor{red}{???The last sentence is unclear???}

@@

To aid in notation, we define a bond matrix $B_{ij}$ such that $B_{ij}=1$ if elements $i$ and $j$ are bonded together and  $B_{ij}=0$ otherwise. Using this notation, the interactive force $(\vec{F}_{IA})_{ij}$ on an element $i$ by an element $j$ is given by
\begin{equation}
\label{F_interactione}
(\vec{F}_{IA})_{ij}= \left\{%
  \begin{array}{ll}
   (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
    & \mbox{if $d_{ij} \leq L_{ij} $ }\\
 %     & \mbox{if $|x_{i} -x_{j}| \leq \frac{1}{2}(D_{i}+D_{j}) $ }\\
    (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=1 $ }\\
       & \mbox{if $d_{ij} >  L_{ij}  \textrm{ and } B_{ij}=1 $ }\\
      0
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=0 $ }.\\
      & \mbox{if $d_{ij} > L_{ij} \textrm{ and } B_{ij}=0 $ }.\\
  \end{array}
\right.
\end{equation}
$(\vec{F_{e}})_{ij}$ and $(\vec{F_{d}})_{ij}$ are the elastic and frictional damping components of the interactive force between elements $i$ and $j$.
The elastic force $(\vec{F_{e}})_{ij}$ between elements is given by
\begin{equation}
\label{F_elastic}
(\vec{F_{e}})_{ij}= -\kappa_{e}  \bigg( d_{i,j}-L_{i,j} \bigg) M_{i,j} \vec{r}_{ij},
\end{equation}
where $\vec{r}_{ij}= \frac{(\vec{x_{i}} -\vec{x_{j}})}{|\vec{x_{i}} -\vec{x_{j}}|}$ is the directional unit vector between the position of element $i$ and $j$, $\kappa_{e}$ is the spring constant, and $M_{i,j}$ is the minimum of the masses of elements $i$ and $j$. The interactive forces obey Newton's 3rd Law (i.e.: $(\vec{F}_{IA})_{ij}=-(\vec{F}_{IA})_{ji}$).
%This is in contrast to some other DEM models in the literature (e.g. \citet{Li2014}). 
The minimum mass, $M_{i,j}$, is preferred to the average thickness, since this means that for two bonded elements a fixed distance apart, the acceleration due to elastic forces is bounded, even when the thickness of one of the elements approaches zero.

@@

The frictional damping force acts to dampen the relative motion of the two particles.
%, both in the direction parallel to $\vec{r}_{ij}$ and in the direction perpendicular to  $\vec{r}_{ij}$.
If $\vec{r}_{ij}^{\perp}$ is the direction vector perpendicular to $\vec{r}_{ij}$, and $P_{\vec{r}_{ij}}$ and $P_{\vec{r}_{ij}^{\perp}}$ are the projection matrices that project onto $\vec{r}_{ij}$ and $\vec{r}_{ij}^{\perp}$ respectively, then the frictional damping force is given by
\begin{equation}
\label{F_damping} 
%(F_{d})_{ij}=   -C_{r}|(u_{i} -u_{j}) \cdot \vec{r}   | \vec{r}  - C_{t}|(u_{i} -u_{j})\cdot  \vec{r}^{\perp} | \vec{r}^{\perp})
(\vec{F_{d}})_{ij}=   \Bigg (-c_{r_{||}} P_{\vec{r}_{ij}}  - c_{r_{\perp}} P_{\vec{r}_{ij}^{\perp}} \Bigg ) \cdot (\vec{u_{i}} -\vec{u_{j}} ) 
\end{equation}
Here $c_{r_{||}}$ and $c_{r_{\perp}}$ are the drag coefficients for the damping motion parallel and perpendicular to $r_{ij}$, respectively. We set $c_{r_{||}}=2\sqrt{\kappa_{e} }$, so that the elastic force parallel to $\vec{r}_{ij}$ is critically damped. The perpendicular drag coefficient is set to $c_{r_{\perp}}=\frac{1}{4}c_{r_{||}}$. The perpendicular damping force is used reduce the relative motion of ice elements passing by one another with overlapping boundaries.
%The perpendicular damping force is used to damping coefficient in the the perpendicular direction is set to $c_{r_{\perp}}=\frac{1}{4}c_{r_{||}}$. 
%The frictional damping force perpendicular to $\vec{r}_{ij}$ helps to reduce the motion of particles when many particles are packed tightly together (as in the uncoupled experiments below), and also acts to reduce the vibrations within larger bonded structures. 
The damping forces are implemented using an implicit time stepping scheme, to avoid stability issues for very small elements (details found in Appendix B).

Figure \ref{fig:Regular_Uncoupled_sim} illustrates the effectiveness of the repulsive forces in an uncoupled (ice-only) simulation of ice elements forced to drift westward into a bay. The elements eventually come to rest in the bay with minimal overlap between elements. 
%The overlap between elements can be reduced by increasing the spring constant, $\kappa_{e}$.
%Since the Courant-Friedrichs-Lewy (CFL) condition for a critically damped spring is $(dt)^{2} < 4 /  \kappa_{e}$, a value of $ \kappa_{e}=10^{-5}$ is used which is large enough to prevent a lot of overlap between elements, but sufficiently small to make the model stable for 5 minute time steps . 
%The overlap of the elements can be reduced by increasing the spring constant, $\kappa_{e}$. A value of $ \kappa_{e}=10^{-5}$ is used which since this is large enough to prevent too much overlap, and small enough to satisfy to the Courant-Friedrichs-Lewy (CFL) stability condition $(dt)^{2} < 4 /  \kappa_{e}$, using a 5 minute time step (see Supporting Information for details).
Figure \ref{fig:Uncoupled_sim} illustrates the effectiveness of the numerical bonds in simulations of small icebergs (individual un-bonded elements) and large icebergs (constructed from many ice elements bonded together) forced to drift towards a convex coast line. When the tabular icebergs arrive at the coast, they bump into the coastline and begin to rotate, influencing the paths of the other icebergs (movies S1 and S2 in the Supporting Information). This example illustrates an advantage of using small elements bonded together, to represent large-scale structure and rotational motion without explicitly accounting for  the angular momentum of the elements (as discussed in \citet{Jakobsen2001}).

\subsection{Initializing element geometry and packing}
For purposes of initialization, we assume that elements have surface areas which are shaped as equally-sized regular hexagons (note that the elements are assumed to be circular for proposes of interactions). When packing elements together, the hexagonal elements are initially arranged in a staggered lattice, with each element bonded to the adjacent elements (Figure \ref{fig:Hex_intersections}a). In this arrangement, each element (away from the edges) is bonded to six other elements. The bonds between elements form a pattern of equilateral triangles, which gives rigidity to the larger structure. The circular shape of elements (used for interactions) is inscribed within the hexagonal shape used for packing (Figure  \ref{fig:Hex_intersections}a). The centers of adjacent elements are initially separated by a distance $d_{i,j}=L_{i,j}=2A_{p}$, where $A_{p}$ is the length the apothems of the hexagons.
% and \ref{fig:Hex_schematic}

@@

\subsection{Ocean-ice and ice-ocean coupling}
The LBIM is coupled to the ocean model via a two-way synchronous coupling, meaning that ocean-model fields are passed to the LBIM, and the LBIM fields are passed back to the ocean model at every time step. Passing fields between the two models involves interpolating the fields from the ocean model's Eularian grid onto the LBIM's `Lagrangian grid' (i.e.: onto the ice elements), and aggregating fields from the Lagrangian elements onto the ocean-model's Eularian grid.

The coupling from the ocean model to the LBIM is straight forward: at every time step: the ocean mixed layer temperature, salinity, velocity and sea-ice concentration are passed from the ocean model to the LBIM, to be used in the momentum and thermodynamic equations of the ice elements. 
Since tabular icebergs are explicitly resolved in the ocean, it is sufficient for each element to interact with ocean mixed layer only (i.e.: there is no need to manually embed icebergs into the ocean by integrating ocean fields over the icebergs' thickness, as suggested in \citet{Merino2016}).
Within the LBIM, the ocean model fields are interpolated onto the Lagrangian grid using a bilinear interpolation scheme.% The LBIM is not sensitive to the chosen interpolation scheme.

%More care needs to be taken when passing fields back from the LBIM to the ocean model. 
%The coupling from the LBIM to ocean model is more complex.
The LBIM influences the ocean by: (i) applying a pressure to the ocean surface, (ii) affecting the upper ocean by applying a no-slip boundary condition and frictional velocity beneath the ice, and (iii) imposing heat, salt and mass fluxes on the ocean, associated with ice melting.
Six fields are passed from the LBIM to the ocean model: ice mass, ice area, frictional velocity, and heat, salt and mass fluxes. 
Fields from the LBIM are aggregated from the Lagrangian elements to the Eulerian ocean grid before they are passed to the ocean model.

The aggregation of the LBIM fields onto the ocean grid is done in a way that is consistent with the shape of the elements in the LBIM (see Section 2.3). Fields are `spread' to the ocean model grid by exactly calculating what fraction of an element's surface area lies in a particular grid box, and dividing the field in proportion to this fraction.
%While the position of a particular element might be inside one ocean grid cell, the implied shape of the element will often intersect other ocean grid cells. 
As an example, consider a hexagonal element in the LBIM, which is positioned such that it intersects four ocean grid cells (Figure \ref{fig:Hex_intersections}b). In this situation, the element's mass (for example) is divided between these four ocean cells in proportion to the overlap area between the hexagonal element and the grid cell (this fraction is shown by the colors in Figure \ref{fig:Hex_intersections}b). An advantage of this approach is that there are no jumps in pressure as an element moves from one grid cell to another, which could trigger artificial tsunamis within the ocean model, making the ocean model unrealistic. 
%This prevents sudden changes to the aggregated pressure field pressure which can trigger tsunamis within the ocean model, making the ocean model unrealistic.  
%The aggregation scheme is therefore designed such that there are no jumps in pressure as an element moves from one grid cell to the next, and total pressure is conserved as an element moves.

The numerical calculation of the intersection between hexagons and the ocean grid is simplified by dividing the hexagon into 6 equilateral triangles. This method allows for the intersection to be found even when the hexagon is not aligned with the grid. %The LBIM can be run using hexagonal or rectangular elements.

@@

When multiple elements are bonded together to form larger structures, it is no longer appropriate to use the parameterizations for melt developed for individual point-particle icebergs. An element which is completely surrounded by other elements, is meant to represent a column of ice in the middle of a large structure, and hence will not experience melt at its sides due to wave erosion or buoyant convection. Also, the iceberg basal melt rate, $M_{b}$ described above is based on boundary layer theory of flow past a finite plate, and is only appropriate for basal surfaces where the distance from the leading edge is sufficiently small \citep{Eckert1950, Weeks1973}. For an element in the interior of a large structure, the distance from the edge of the structure is large, and so using $M_{b}$ for the basal melt is not appropriate. Instead, the basal melt, $M_{s}$ is determined using the three equation model for basal melt, which is a typical melting parametrization used beneath ice shelves \citep{Holland1999}.% (see Appendix A).

@@

\subsubsection{Parallelization and halos}
The LBIM runs on multiple processors in parallel (using the same grid decomposition as the ocean model). When elements move from an ocean cell on one processor to an ocean cell on a second processor, the memory has to be passed from one processor the next, added and removed to the appropriate lists and the memory has to be allocated and deallocated correctly. Element interactions across the edge of processors are handled using computational halos. A computational halo is a copy of the edge of a one processor which is appended to the edge of a second processor, so that the first processor can interact with the second processor during a time step. Before each time step, elements at the edges of each processor are copied onto the halos of adjacent processors so that they can be used in calculating the interactive forces. After each time step, these halos are emptied, and the process is repeated. These halo updates are one of the most computationally expensive parts of the LBIM. Details of how the bonds are broken and reconnected across processor boundaries are provided in Appendix~C.

>>>>

\authors{A. A. Stern\affil{1}, A. Adcroft\affil{1}, O. Sergienko\affil{1}, G. Marques\affil{1}}

@@

\begin{abstract}
Large tabular icebergs calved from Antarctic ice shelves have long lifetimes (due to their large size), during which they drift across large distances, altering ambient ocean circulation, bottom-water formation, sea-ice formation and biological primary productivity in the icebergs' vicinity.
However, despite their importance, the current generation of ocean circulation models do not represent large tabular icebergs.
In this study we develop a novel framework to model large tabular icebergs submerged in the ocean. In this framework, tabular icebergs are represented by Lagrangian elements that drift in the ocean, and are held together and interact with each other via bonds.
%We describe the Lagrangian bonded iceberg model and discuss a number of the technical elements involved in its development. 
%We demonstrate the capabilities of the model by modeling a tabular iceberg drifting away an idealized ice shelf.} 
A break of these bonds allows the model to emulate calving events (i.e. detachment of a tabular iceberg from an ice shelf) and tabular icebergs breaking up into smaller pieces. Idealized simulations of a calving tabular iceberg, its drift, and its breakup, demonstrate capabilities of the developed framework.
\end{abstract}

@@

In order to allow icebergs to travel large distances, we model the icebergs in a Lagrangian framework (as in the point particle iceberg drift models described above). However in our model icebergs are no longer treated as point particles that interact with the ocean at a single location. Instead icebergs are given physical structure, so that they interact with the ocean across multiple ocean grid cells, depress the ocean surface over a wide area, and can interact with other icebergs (Figure \ref{fig:Hex_schematic}). This is done by assigning a finite surface area and shape to the Lagrangian elements, which allows the elements to behave as if they have a finite extent. 
%Element interact with the ocean by distributing their weight, surface area and melt fluxes over multiple ocean grid cells in a way which is consistent with the shape of the ice element.
The finite extent of an element is transmitted by the ocean by distributing the element's weight, surface area and melt fluxes over multiple ocean grid cells in a way which is consistent with the shape of the ice element. 
Finite-extent elements interact with each other via repulsive forces which are applied when the boundaries of the elements overlap. This prevents the icebergs from piling up on top of one another, which has been an issue near coastlines in previous point-particle icebergs models.
%Elements interact with other elements when the boundaries of the elements overlap (see schematic in Figure \ref{fig:Hex_schematic}).

Large tabular icebergs can then be represented by `bonding' together multiple ice elements into larger structures using numerical bonds (Figure \ref{fig:Hex_schematic}).
The numerical bonds hold the ice elements together and allow a collection of elements to move as a unit. 
This allows tabular icebergs to drift in the ocean when forced by ocean currents and wind.
%Figure \ref{fig:Hex_schematic}) shows a schematic of the difference between point-particle icebergs, finite extent icebergs and tabular icebergs constructed out of multiple elements. 
%Figure \ref{fig:Hex_schematic} shows a schematic of point-particle elements, elements with finite extent and bonded elements, and shows schematically how tabular icebergs can be constructed out of multiple elements.
An advantage of representing tabular icebergs using numerical bonds is that by breaking the bonds, we can simulate iceberg calving (e.g.: Figure \ref{fig:SST_Collapse}), or the response to an iceberg fracturing into multiple smaller pieces (see movies S1 and S2 in the Supporting Information).

@@

%As mentioned above, we simplify the KID model by choosing to not explicitly calculate the angular momentum of the ice elements, which is computationally expensive. Instead, we allow rotational motion of larger structures emerge as a consequence of bond orientation and collective motion \citep{Jakobsen2001}.
%

@@

The Kinematic Iceberg Dynamics model (KID) is a Lagrangian particle-based model in that the objects of the model are Lagrangian elements. Each element represents a column of ice that is floating in the ocean, and has a position, velocity, mass, and a set of dimensions, which can evolve in time. The motion of each element is determined by a momentum equation which is solved in the (Lagrangian) reference frame of the element. The elements experience oceanic and atmospheric forces, which are either prescribed, or computed by coupling the iceberg model to an ocean/atmosphere model. The ice elements also interact with one another via attractive and repulsive interactive forces, and can be bonded together to form larger structures. The angular momentum of the elements is not modeled explicitly; instead rotational motion of larger structures emerge as a consequence of bond orientation and collective motion.

In different contexts, the ice elements can be thought to represent individual icebergs, sea ice flows, or, when the elements are bonded together, they can represent larger structures such as tabular icebergs or ice shelves.

%In this latter case, the KID can be categorized as a discrete element model (DEM). 
% in the sense that large structures are constructed from smaller independent Lagrangian elements.

The KID model is developed on the code base of an existing iceberg drift model \citep{Martin2010, Stern2016}. When run with the correct set of runtime flags, the model runs as a traditional point-particle iceberg drift model.
%
%In this section, we discuss the equations of motion governing the element movement and the thermodynamic equations governing the decay of the elements. The also discuss the interactive forces

@@

% 
When ice elements move alone (without interactions with other elements), they can be thought of as representing individual (or clusters of) small icebergs, and follow the same equations described in the iceberg drift model of \citet{Martin2010} (based on the equations outlined in \citet{Bigg1997} and \citet{Gladstone2001}).
A description of these forces is provided for completeness in Appendix A.

In addition to the external forces, the ice elements experience interactive forces due to the presence of other elements. Two types of interactive forces are included between elements. The first force is a repulsive force which is applied to elements to prevent them from overlapping the boundaries of the neighboring elements. The second interactive force is a force due to numerical `bonds', and is only applied if two elements are labelled as `bonded'. When two elements are bonded, each element feels an attractive force that prevents the elements from moving too far apart from one another.  The details of the interactive forces are provided in below. 
%The interactive forces between two bonded elements are defined such that in the absence of other forces the elements come to rest adjacent to one another, with no overlap of the iceberg areas. \textcolor{red}{???The last sentence is unclear???}

@@

To aid in notation, we define a bond matrix $B_{ij}$ such that $B_{ij}=1$ if elements $i$ and $j$ are bonded together and  $B_{ij}=0$ otherwise. Using this notation, the interactive force $(\vec{F}_{IA})_{ij}$ on an element $i$ by an element $j$ is given by
\begin{equation}
\label{F_interactione}
(\vec{F}_{IA})_{ij}= \left\{%
  \begin{array}{ll}
   (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
    & \mbox{if $d_{ij} \leq L_{ij} $ }\\
 %     & \mbox{if $|x_{i} -x_{j}| \leq \frac{1}{2}(D_{i}+D_{j}) $ }\\
    (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=1 $ }\\
       & \mbox{if $d_{ij} >  L_{ij}  \textrm{ and } B_{ij}=1 $ }\\
      0
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=0 $ }.\\
      & \mbox{if $d_{ij} > L_{ij} \textrm{ and } B_{ij}=0 $ }.\\
  \end{array}
\right.
\end{equation}
$(\vec{F_{e}})_{ij}$ and $(\vec{F_{d}})_{ij}$ are the elastic and frictional damping components of the interactive force between elements $i$ and $j$.
The elastic force $(\vec{F_{e}})_{ij}$ between elements is given by
\begin{equation}
\label{F_elastic}
(\vec{F_{e}})_{ij}= -\kappa_{e}  \bigg( d_{i,j}-L_{i,j} \bigg) M_{i,j} \vec{r}_{ij},
\end{equation}
where $\vec{r}_{ij}= \frac{(\vec{x_{i}} -\vec{x_{j}})}{|\vec{x_{i}} -\vec{x_{j}}|}$ is the directional unit vector between the position of element $i$ and $j$, $\kappa_{e}$ is the spring constant, and $M_{i,j}$ is the minimum of the masses of elements $i$ and $j$. The interactive forces obey Newton's 3rd Law (i.e.: $(\vec{F}_{IA})_{ij}=-(\vec{F}_{IA})_{ji}$).
%This is in contrast to some other DEM models in the literature (e.g. \citet{Li2014}). 
The minimum mass, $M_{i,j}$, is preferred to the average mass, since this means that for two bonded elements a fixed distance apart, the acceleration due to elastic forces is bounded, even when the mass of one of the elements approaches zero.

@@

The frictional damping force acts to dampen the relative motion of the two particles.
%, both in the direction parallel to $\vec{r}_{ij}$ and in the direction perpendicular to  $\vec{r}_{ij}$.
If $\vec{r}_{ij}^{\perp}$ is the direction vector perpendicular to $\vec{r}_{ij}$, and $P_{\vec{r}_{ij}}$ and $P_{\vec{r}_{ij}^{\perp}}$ are the projection matrices that project onto $\vec{r}_{ij}$ and $\vec{r}_{ij}^{\perp}$ respectively, then the frictional damping force is given by
\begin{equation}
\label{F_damping} 
%(F_{d})_{ij}=   -C_{r}|(u_{i} -u_{j}) \cdot \vec{r}   | \vec{r}  - C_{t}|(u_{i} -u_{j})\cdot  \vec{r}^{\perp} | \vec{r}^{\perp})
(\vec{F_{d}})_{ij}=   \Bigg (-c_{r_{||}} P_{\vec{r}_{ij}}  - c_{r_{\perp}} P_{\vec{r}_{ij}^{\perp}} \Bigg ) \cdot (\vec{u_{i}} -\vec{u_{j}} ) 
\end{equation}
Here $c_{r_{||}}$ and $c_{r_{\perp}}$ are the drag coefficients for the damping motion parallel and perpendicular to $r_{ij}$, respectively. We set $c_{r_{||}}=2\sqrt{\kappa_{e} }$, so that the elastic force parallel to $\vec{r}_{ij}$ is critically damped. The perpendicular drag coefficient is set to $c_{r_{\perp}}=\frac{1}{4}c_{r_{||}}$. The perpendicular damping force is used reduce the relative motion of ice elements passing by one another with overlapping boundaries.
%The perpendicular damping force is used to damping coefficient in the the perpendicular direction is set to $c_{r_{\perp}}=\frac{1}{4}c_{r_{||}}$. 
%The frictional damping force perpendicular to $\vec{r}_{ij}$ helps to reduce the motion of particles when many particles are packed tightly together (as in the uncoupled experiments below), and also acts to reduce the vibrations within larger bonded structures. 
The damping forces are implemented using an implicit time stepping scheme, to avoid stability issues for very small elements (details found in Appendix B). %The system is numerically stable for time steps $(dt)^{2} < 4 /  \kappa_{e}$.
% In the examples below, we use a spring constant $ \kappa_{e}=10^{-5}$, which allows the model to be stable using a 10 minute time step (although smaller time steps are often used).

Figure \ref{fig:Regular_Uncoupled_sim} illustrates the effectiveness of the repulsive forces in an uncoupled (ice-only) simulation. In this simulation ice elements are forced westward into a bay, and eventually come to rest in the bay with a small amount of overlap between elements.
The amount of overlap between elements in the final state of the simulation depends on the magnitude of the spring constant, $\kappa_{e}$, with larger spring constants reducing the amount of overlap. Increasing the spring constant also makes the system numerically stiff so that smaller time steps are required to prevent numerical instabilities (the system is stable for time steps satisfying $dt^{2} < 4 /  \kappa_{e}$).  A value of $ \kappa_{e}=10^{-5}$ is chosen that is large enough to prevent too much overlap between elements for typical ocean forcings (e.g: Figure \ref{fig:Regular_Uncoupled_sim}), and small enough to allow for time steps up to 10 minutes (smaller time steps are used when the model is coupled to an ocean model). %(although smaller time steps are often used to prevent the ocean model crashing or introducing artificial effects gravity waves).

Figure \ref{fig:Uncoupled_sim} illustrates the effectiveness of the numerical bonds in simulations of small icebergs (individual un-bonded elements) and large icebergs (constructed from many ice elements bonded together) forced to drift towards a convex coast line. When the tabular icebergs arrive at the coast, they bump into the coastline and begin to rotate, influencing the paths of the other icebergs. This example illustrates an advantage of using small elements bonded together to represent large-scale structure - i.e. rotational motion of large structures can be simulated without explicitly accounting for the angular momentum of the elements (as discussed in \citet{Jakobsen2001}). Movies of these uncoupled simulations are found in S3 and S4 in the Supporting Information.

\subsection{Initializing element geometry and packing}
For purposes of initialization, we assume that elements have surface areas which are shaped as equally-sized regular hexagons (note that the elements are assumed to be circular for proposes of interactions). When packing elements together, the hexagonal elements are initially arranged in a staggered lattice, with each element bonded to the adjacent elements (Figures \ref{fig:Hex_schematic} and \ref{fig:Hex_intersections}a). In this arrangement, each element (away from the edges) is bonded to six other elements. The bonds between elements form a pattern of equilateral triangles, which gives rigidity to the larger structure. The circular shape of elements (used for interactions) is inscribed within the hexagonal shape used for packing (Figure  \ref{fig:Hex_intersections}a). The centers of adjacent elements are initially separated by a distance $d_{i,j}=L_{i,j}=2A_{p}$, where $A_{p}$ is the length the apothems of the hexagons.
% and \ref{fig:Hex_schematic}

@@

\subsection{Ocean-ice and ice-ocean coupling}
The KID model is coupled to the ocean model via a two-way synchronous coupling, meaning that ocean-model fields are passed to the iceberg model, and iceberg model fields are passed back to the ocean model at every time step. Passing fields between the two models involves interpolating the fields from the ocean model's Eularian grid onto the iceberg model's `Lagrangian grid' (i.e.: onto the ice elements), and aggregating fields from the Lagrangian elements onto the ocean-model's Eularian grid.

The coupling from the ocean model to the iceberg model is straight forward: at every time step: the ocean mixed layer temperature, salinity, velocity and sea-ice concentration are passed from the ocean model to the iceberg model, to be used in the momentum and thermodynamic equations of the ice elements. 
Since tabular icebergs are explicitly resolved in the ocean, it is sufficient for each element to interact with ocean mixed layer only (i.e.: there is no need to manually embed icebergs into the ocean by integrating ocean fields over the icebergs' thickness, as suggested in \citet{Merino2016}).
Within the KID model, the ocean model fields are interpolated onto the Lagrangian grid using a bilinear interpolation scheme.% The KID is not sensitive to the chosen interpolation scheme.

%More care needs to be taken when passing fields back from the KID to the ocean model. 
%The coupling from the KID to ocean model is more complex.
The iceberg model influences the ocean by: (i) applying a pressure to the ocean surface, (ii) affecting the upper ocean by applying a no-slip boundary condition and frictional velocity beneath the ice, and (iii) imposing heat, salt and mass fluxes on the ocean, associated with ice melting.
Six fields are passed from the iceberg model to the ocean model: ice mass, ice area, frictional velocity, and heat, salt and mass fluxes. 
Fields in the iceberg model are aggregated from the Lagrangian elements to the Eulerian ocean grid before they are passed to the ocean model.

The aggregation of the iceberg-model fields onto the ocean grid is done in a way that is consistent with the shape of the elements in the iceberg model (see Section 2.3). Fields are `spread' to the ocean model grid by exactly calculating what fraction of an element's surface area lies in a particular grid box, and dividing the field in proportion to this fraction.
%While the position of a particular element might be inside one ocean grid cell, the implied shape of the element will often intersect other ocean grid cells. 
As an example, consider a hexagonal element in the iceberg model, which is positioned such that it intersects four ocean grid cells (Figure \ref{fig:Hex_intersections}b). In this situation, the element's mass (for example) is divided between these four ocean cells in proportion to the overlap area between the hexagonal element and the grid cell (this fraction is shown by the colors in Figure \ref{fig:Hex_intersections}b). An advantage of this approach is that there are no jumps in pressure as an element moves from one grid cell to another, which could trigger artificial tsunamis within the ocean model, making the ocean model unrealistic. 
%This prevents sudden changes to the aggregated pressure field pressure which can trigger tsunamis within the ocean model, making the ocean model unrealistic.  
%The aggregation scheme is therefore designed such that there are no jumps in pressure as an element moves from one grid cell to the next, and total pressure is conserved as an element moves.

The numerical calculation of the intersection between hexagons and the ocean grid is simplified by dividing the hexagon into 6 equilateral triangles. This method allows for the intersection to be found even when the hexagon is not aligned with the grid. %The KID can be run using hexagonal or rectangular elements.

@@

When multiple elements are bonded together to form larger structures, it is no longer appropriate to use the melt parameterizations developed for individual point-particle icebergs. An element which is completely surrounded by other elements, is meant to represent a column of ice in the middle of a large structure, and hence will not experience melt at its sides due to wave erosion or buoyant convection. Also, the iceberg basal melt rate, $M_{b}$ described above is based on boundary layer theory of flow past a finite plate, and is only appropriate for basal surfaces where the distance from the leading edge is sufficiently small \citep{Eckert1950, Weeks1973}. For an element in the interior of a large structure, the distance from the edge of the structure is large, and so using $M_{b}$ for the basal melt is not appropriate. Instead, the basal melt, $M_{s}$ is determined using the three equation model for basal melt, which is a typical melting parametrization used beneath ice shelves \citep{Holland1999}.% (see Appendix A).

@@

\subsubsection{Parallelization and halos}
The KID model runs on multiple processors in parallel (using the same grid decomposition as the ocean model). When elements move from an ocean cell on one processor to an ocean cell on a second processor, the memory has to be passed from one processor the next, added and removed to the appropriate lists and the memory has to be allocated and deallocated correctly. Element interactions across the edge of processors are handled using computational halos. A computational halo is a copy of the edge of a one processor which is appended to the edge of a second processor, so that the first processor can interact with the second processor during a time step. Before each time step, elements at the edges of each processor are copied onto the halos of adjacent processors so that they can be used in calculating the interactive forces. After each time step, these halos are emptied, and the process is repeated. These halo updates are one of the most computationally expensive parts of the iceberg model. Details of how the bonds are broken and reconnected across processor boundaries are provided in Appendix~C.
